Individual heterogeneity in the functional network topography of the medial prefrontal cortex

### 2019-05-03 Manuscript.Rmd
discussion draft done

<!--
We also diagnosed our choice of two communities by reporting the rank (from lowest to largest) of the first eigenvalue produced by SP to be non-zero. This eigenvalue serves as an estimator of the optimal number of communities in a network (Chung 1999), such that a rank of two in our partitions justifies limiting the number of communities to two.
-->

@@

We performed a number of comparisons among partitions. First, we computed the degree of agreement between SP and modularity per subject. SP and Modularity have been previously found to show underfitting and overfitting tendencies, respectively, in their community detection performance in a diverse number of network types [@Ghasemian2018], so alignment between the two algorithms would increase our confidence in the validity of the resulting partitionings. Next, we compared the subject-level SP partitionings across individuals, and calculated the mean pairwise ARI for the group. We then performed the same evaluation for PCC and mPFC separately, and examined whether there were mean differences in overall agreement within these regions by performing a paired permutation analysis (5000 iterations) on all pairwise ARI comparisons across subjects. Given that the functional heterogeneity of brain areas is unevenly distributed across the cortex [@Mueller2013a].

@@

# # eigenvalue-based diagnosis of optimal n of communities
# prcntDissection <- lapply(Summaries, function(data) {which(data$EigenVal < 0.00001)}) # how many consecutive 0s?
# prcntDissection <- mean((sapply(prcntDissection, length) + 1) == 2) * nSubj # % of subjects who only had 1 zero (thus optimal cuts = 2)

@@

The two clustering methods had high agreement (mean ARI = `r round(mean(RI_all_within$Index), digits = 2)`, SD = `r round(sd(RI_all_within$Index), digits = 2)`). Modularity showed a tendency to produce additional communities (median = `r median(nModularity)`, range = `r range(nModularity)`). However, the additional communities encompassed a small number of vertices (median = `r median(modularityCounts_extra)`, IQR = `r quantile(modularityCounts_extra, 0.25)` - `r quantile(modularityCounts_extra, 0.75)`) compared to the principal two (median = `r median(modularityCounts_two)`, IQR = `r quantile(modularityCounts_two, 0.25)` - `r quantile(modularityCounts_two, 0.75)`), suggesting that a binary partitioning could provide a reasonable approximation of the network's true community structure. <!--  In addition, the estimation of the optimal number of cuts based on the rank of the first non-zero eigenvalue (see "Partition evaluation" for details) showed that 100 percent of the individual networks were best partitioned into two communities. -->

@@

The thresholded partitions in our work qualitatively resembled the topographical arrangement of DN sub-networks A and B proposed by Braga and Buckner [-@Braga2017]. We explored this similarity by reproducing their seed-based connectivity approach in three of our subjects. In their original work, Braga and Buckner manually selected individual vertices in areas including the dorsolateral prefrontal cortex (DLPFC), temporo-parietal junction (TPJ), and parahippocampal cortex (PHC) that produced two spatially anticorrelated networks [@Braga2017; @Braga2019]. Therefore, we hypothesized that if the communities we identified on the medial wall matched their proposed networks, we should be able to reproduce our partitionings by selecting seeds in these diagnostic areas. For functional connectomes from two individuals (100307 and 101006), we selected seeds for networks A and B in TPJ (seed indices), and confirmed their placement based on functional connectivity patterns observed in TPJ, PCC, DLPFC, and PHC. The whole-brain seed-based functional connectivity maps for two individuals are juxtaposed with the corresponding community detection results in Figure 8.

Visual inspection of these networks shows a remarkable similarity between our DN community and the previously reported sub-network A. However, the non-DN community filled areas not covered by either DN-A or DN-B, further evidencing its disentanglement from DN processes. Subthreshold vertices from the DN community overlaped with DN-B vertices, suggesting that this network could act as a dynamic intermediary between DN-A and other networks.

Understanding the interaction of these three networks will require future examination. Nevertheless, these qualitative similarities are important, as they collectively suggest that: (1) canonical DN regions can be topographically fractionated into DN and non-DN; (2) the DN community can be further partitioned into two sub-networks; (3) the SP algorithm is a useful and reliable tool for identifying this refined topography; and (4) the topographic organization of the communities is variable across a large sample of participants, especially in mPFC.

@@

Here we address these challenges by using spectral partitioning (SP), a graph-theoretic community detection algorithm that efficiently separates a network into two [@Fiedler1975; @Higham2007; @Toker2019]. For 100 individuals, we reliably subdivided brain regions that typically show overlap between DN and valuation effects into DN and non-DN communities. Restricting our analyses to a general mPFC/PCC search space made it appropriate to use a technique that identified a vertex-wise, binary partitioning that was sensitive to the complex topography of the brain. This contrasts with whole-brain network analyses, which need to allow for multiple sub-networks and which often use coarser-grained parcels as the units of analysis. Partitioning an individual's brain network through SP has a number of advantages, such as identifying communities deterministically, constraining comunities to contain a similar number of vertices (i.e. preventing the allocation of most vertices to a single community), providing continuous values to determine the strength of a node's community affiliation, and diagnosing the connectedness and separability of a network through examination of its resulting eigenvalues (**TODO: show this in text**) [@chung1997; @Higham2007]. Comparisons with partitionings formed by modularity maximization, which heuristically determines the ideal number of communities, suggested the binary partitioning was appropriate.

@@

Our results are complementary with previous studies that described individual-specific brain organization. On one hand, topographical heterogeneity has been qulitatively highlighted using a different data aspect ratio [few individuals and more data per individual; @Gordon2017, @Braga2017, @Braga2019]. In terms of stability, previous work has relied on parcellations and group priors to show that an individual's functional connectivity pattern is consistent in time [@Gratton2018; @Kong2018], thus being relatively insensitive to an individual's complex topography. Here we were able to quantify the variability and stability of fine-scale functional topography in a large sample with moderately low amounts of data (down to a single 15 minute scan). The motivation to sub-divide DN also stems from recent work by Kernbach et al. [-@Kernbach2018], where they identified specialized communication of parcels within DN with the rest of the brain in a large pool of individuals.

We found SP results could be made more stable by thresholding the underlying Fiedler vector. We generated custom thresholds for each individual by interrogating the relationship between Fiedler vector values and the degree of persistent community affiliation of vertices over time. The magnitude of these eigenvector values has been recently used to characterize the continuous connectivity profile of the insula with the rest of the brain, challenging the notion of discrete parcellations of this region [@Tian2018]. We anticipate that these values will be helpful to future neuroscientific studies in providing gradual information not afforded by other clustering algorithms.

Outline:

  - SP outperformed seed-based correlation. This makes sense, since it makes use of more information (n(n-1)/2 pairwise correlations, rather than n correlations with the seed).

  - Our work is complementary with Braga & Buckner. They identified subdivisions, whereas we've identified variable topography for the DN as a whole.

  - Limitations. Our approach is (by design) more anatomically focal. A binary partitioning is undoubtedly only an approximation of the ground truth. Etc.

  - Future directions. This could be used to identify individualized units of analysis (ROIs) for task activation studies. This might be useful in clinical studies, or in studies aiming to dissociate different forms of valuation, or to reach a stronger conclusion as to whether valuation and DMN processes are topographically segregable or colocalized.

<!--
- We are also not the first to show functional subdivision of DN [@Kernbach2018] (but that's using group averages on parcellated data), or this particular individualized pattern of subdivision (braga). Our results are complementary in the sense that we show similar outcomes in 100 subjects. The major discrepancy is that we assign meaning to the communities based on their coverage of regions identified through a meta-analysis.
- Stability within subject has also been shown before. This work was done using either parcellations and oversampling though. [@Laumann2015; @Gordon2017; @Braga2017; @Gratton2018b; @Kong2018]
- Talk about SP as a way to identify an individual's partition: deterministic, with simple diagnostics, and requires less data requirements.
- Similarity with B&B: One way in which our results are at odds with those from Braga and Buckner is the functional labeling of these sub-networks. Based on our criteria, network B would more closely identify with DN (per 7m affiliation), while we currently do not have enough evidence to label network A. These discrepancies should be addressed with explicit behavioral manipulations in the future. Also our search space is limited compared to theirs, thus making the patterns only partially similar.
- The main gap is mapping behavior to functional organization, but we will do that with behavioral manipulations of subjective value in the future. Instead, we focus on understanding the importance of individualized mapping in these areas first, as a baseline.
- Clinical implications: individualized targets, and more accurate estimation of impacts based on assessment of individual patients. Similarly, our approach focuses on individual estimations without group priors, so it's well fit to address the problem of low n in clinical research. Also, SP can produce these effects with moderate amounts of data. Previous results suggest impractical amounts.

For clinical implications, see Hiser & Koenigs 2018. In short, being able to identify functional idiosyncrasies without much data can yield important insights in diagnosis (meh, we always say that..) of a variety of psychological diseases associated with mPFC. It would be more interesting to identify the repercussions of damage in mPFC per individual. They mention that subgenual vmPFC is a common target for deep brain stimulation (or TMS) as a depression treatment. The refinement of the technique is thus coupled to being able to identify personalized functional topologies. They list a number of results related to mPFC functional connectivity too. In the end, their suggestions for clinical implications of mPFC studies rely on the individual measure of functional expanse. (note that social-relevant areas in their meta analysis look similar to our patterns, and the emotion part to Barbas' latest)

It will also be good to note that there is concurrent debate on the ground truth of community detection. In other fields, identifying a community structure is not necessarily associated with the true organization of the network. What happens is that you might find a tangential grouping that gives you information different from your original goal [@Clauset2004]. In our case, however, averaging everyone's partitions produces organizational maps that resemble those from previous work (e.g. Margulies).

Talk about how there are other community detection algorithms, and that modularity and SP over and underfit, respectively. However, SP eigenmaps have been found to provide relevant functional information. In addition, both modularity and SP rely on connectivity strength, and thus is amenable to the correlative nature of functional connectivity.

And write a portion on the lack of subcortical reward-related regions (i.e. striatum).

Make sure to emphasize that the high overlap between sliding windows makes it more likely that differences are due to noise in the signal, and not strong variability. Like, you wouldn't expect the actual network distribution to shift that much after shifting the time series by 1 minute.
-->

>>>>

@@

We performed a number of comparisons among partitions. First, we computed the degree of agreement between SP and modularity per subject. SP and Modularity have been previously found to show underfitting and overfitting tendencies, respectively, in their community detection performance in a diverse number of network types [@Ghasemian2018], so alignment between the two algorithms would increase our confidence in the validity of the resulting partitionings. Next, we compared the subject-level SP partitionings across individuals, and calculated the mean pairwise ARI for the group. We then performed the same evaluation for PCC and mPFC separately, and examined whether there were mean differences in overall agreement within these regions by performing a paired permutation analysis (5000 iterations) on all pairwise ARI comparisons across subjects. Given that the functional heterogeneity of brain areas is unevenly distributed across the cortex [@Mueller2013].

@@

@@

The two clustering methods had high agreement (mean ARI = `r round(mean(RI_all_within$Index), digits = 2)`, SD = `r round(sd(RI_all_within$Index), digits = 2)`). Modularity showed a tendency to produce additional communities (median = `r median(nModularity)`, range = `r range(nModularity)`). However, the additional communities encompassed a small number of vertices (median = `r median(modularityCounts_extra)`, IQR = `r quantile(modularityCounts_extra, 0.25)` - `r quantile(modularityCounts_extra, 0.75)`) compared to the principal two (median = `r median(modularityCounts_two)`, IQR = `r quantile(modularityCounts_two, 0.25)` - `r quantile(modularityCounts_two, 0.75)`), suggesting that a binary partitioning could provide a reasonable approximation of the network's true community structure.

@@

The thresholded partitions in our work qualitatively resembled the topographical arrangement of DN sub-networks A and B proposed by Braga and Buckner [-@Braga2017]. We explored this similarity by reproducing their seed-based connectivity approach in three of our subjects. In their original work, Braga and Buckner manually selected individual vertices in areas including the dorsolateral prefrontal cortex (DLPFC), temporo-parietal junction (TPJ), and parahippocampal cortex (PHC) that produced two spatially anticorrelated networks [@Braga2017; @Braga2019]. Therefore, we hypothesized that if the SP communities corresponded their proposed networks, our partitionings should match networks A and B generated by seed-based functional connectivity in these diagnostic areas. For functional connectomes from two individuals (100307 and 101006), we selected seeds for networks A and B in TPJ, and confirmed their placement based on functional connectivity patterns observed in TPJ, PCC, DLPFC, and PHC (correlation coefficients thesholded at 0.2), as stipulated by Braga et al. [-@Braga2019]. The whole-brain seed-based functional connectivity maps for two individuals are juxtaposed with the corresponding community detection results in Figure 8. Visual inspection of these networks shows a remarkable similarity between our DN community and the previously reported sub-network A. However, the non-DN community filled areas not covered by either DN-A or DN-B, further evidencing its disentanglement from DN processes.

@@

Here we address these challenges by using spectral partitioning (SP), a graph-theoretic community detection algorithm that efficiently separates a network into two [@Fiedler1975; @Higham2007; @Toker2019]. For 100 individuals, we reliably subdivided brain regions that typically show overlap between DN and valuation effects into DN and non-DN communities. Restricting our analyses to a general mPFC/PCC search space made it appropriate to use a technique that identified a vertex-wise, binary partitioning that was sensitive to the complex topography of the brain. This contrasts with whole-brain network analyses, which need to allow for multiple sub-networks and which often use coarser-grained parcels as the units of analysis. Partitioning an individual's brain network through SP has a number of advantages, including identifying communities deterministically, constraining comunities to contain a similar number of vertices (i.e. preventing the allocation of most vertices to a single community), providing continuous values to determine the strength of a node's community affiliation, and diagnosing the connectedness of a network through examination of its resulting eigenvalues [@chung1997; @Higham2007]. Comparisons with partitionings formed by modularity maximization, which heuristically determines the ideal number of communities [@Garcia2017], suggested the binary partitioning was appropriate.

@@

Our results are complementary with previous studies that described individual-specific brain organization. On one hand, topographical heterogeneity has been qulitatively highlighted using a different data aspect ratio [few individuals and more data per individual; @Gordon2017, @Braga2017, @Braga2019]. In terms of stability, previous work has relied on parcellations and group priors to show that an individual's functional connectivity pattern is consistent in time [@Gratton2018b; @Kong2018], thus being relatively insensitive to an individual's complex topography. Here we were able to quantify the variability and stability of fine-scale functional topography in a large sample with moderately low amounts of data (down to a single 15 minute scan). The motivation to sub-divide DN also stems from recent work by Kernbach et al. [-@Kernbach2018], where they identified specialized communication of parcels within DN with the rest of the brain in a large pool of individuals.

In addition to its general diagnostics, SP provides analytical advantages specific to neuroscience. We found that SP outperformed a traditional seed-based correlation approach in capturing idiosyncratic functional topography. By relying on all sources of functional covariation among cortical vertices (rather than individual correlations with the seed), community detection methods in general, and SP in particular, are capable of further filtering nodes with weak sub-network affiliation. In addition, we found SP results could be made more stable by thresholding the underlying Fiedler vector by using temporal stability to apply a custom threshold per individual. The magnitude of these eigenvector values has been recently used to characterize the continuous connectivity profile of the insula with the rest of the brain, challenging the notion of discrete parcellations of this region [@Tian2018]. The combination of discrete classification and gradual information provided by SP points to flexible information benefits not regularly afforded by other clustering algorithms.

The community organization of PCC and mPFC shares significant similarities with DN sub-networks A and B proposed by Braga & Buckner [-@Braga2017; -@Braga2019]. The topography of our thresholded DN community closely matched network A, while non-DN captured zones uncovered by either DN network. Subthreshold vertices from the DN community overlaped with DN-B vertices, suggesting that this network could act as a dynamic intermediary between DN-A and other networks. Our findings therefore complement the initial identification of sub-DN networks by quantifying the systematic variability of its underlying topography in a larger group of people Understanding the interaction of networks A (DN), B, and non-DN should be subject of future examinations. Nevertheless, these qualitative similarities are important, as they collectively suggest that: (1) canonical DN regions can be topographically fractionated into DN and non-DN; (2) the DN community can be further partitioned into two sub-networks; (3) the SP algorithm is a useful and reliable tool for identifying this refined topography in a data-driven way; and (4) the topographic organization of the communities is variable across a large sample of participants, especially in mPFC.

The work presented here faces a number of challenges. A prominent limitation is that our approach is (by design) anatomically focal. However, similar spectral methods have been successfully recruited in larger cortical zones. For example, Margulies et al. [-@Margulies2016] used a combination of gradient maps to characterize the spatial embedding of unimodal and heteromodal regions of cortex. Here we show that a variant of such methods, in combination with prior information on the brain space of interest, is capable of producing refined functional topography in individuals. A similar limitation corresponds to forcing a bipartition of this region. While we provide evidence that binarizing this region is adequate, questions remain regarding the potential dynamics that might obscure additional networks (e.g. the correspondence of subthreshold DN vertices to network B, as mentioned above). In addition, structural connections from these networks to the rest of the brain might produce more nuanced information about sub-specialization of these communities [@Kernbach2018].

Our findings pose interesting avenues for future work. The ability to capture an individual's functional topography without the need of group priors has direct implications for clinical research and diagnosis, where acquiring large sample and data sizes are challenging. Studies interested in characterizing the repercussions of mental illness on neural function could benefit from answering this question in individual patients. This is especially true for disorders such as schizophrenia and depression that significantly affect mPFC [@Hiser2018]. mPFC has also recently become a common target for deep brain stimulation (or TMS) as a depression treatment [CITATION FROM HISER AND KOENIGS], but signal quality issues make imaging this area challenging [CITATION]. Being able to produce stable personalized targets in mPFC has the potential to boost our ability to implement and evaluate the effectiveness of these methods. Finally, the main motivating factor for this study was to assess whether cortical regions subserving DN and valuation are indeed overlapping in mPFC and PCC. While our current results suggest a re-examination of this question at the individual level, we do not estimate the location of valuation relative to DN. A natural next step is to examine whether SP can be similarly used to capture the idiosyncratic cortical topography of valuation. This could allow us not only to more strongly conclude whether DN and valuation are neurally segregable, but also aid in studies aiming to dissociate different forms of valuation.
